\begin{abstract}
Hyperspectral image (HSI) classification often suffers from performance degradation under geographical domain shifts due to variations in atmospheric and sensing conditions. Although domain generalization (DG) aims to learn domain-invariant representations without target domain access, existing DG methods often rely primarily on visual representations and lack explicit semantic guidance for robust cross-domain representation learning. Furthermore, progressive generation models remain limited in capturing transferable semantic relationships across domains, which can lead to semantic drift during progressive band synthesis. To address these limitations, we propose the Semantic-Guided Progressive Domain Generalization framework. Our method integrates class-level descriptions into a Prompt Bank using a frozen Contrastive Language-Image Pretraining (CLIP) text encoder. At each stage of the progressive spectral downsampling cascade, a Semantic-Aware Progressive Generator employs a Lightweight Image Encoder and a Multi-Head Cross Attention layer to perform semantic fusion between spatial features and prompt embeddings. The intermediate spatial-spectral representations are modulated via 3D Adaptive Layer Normalization (AdaLN3d) conditioned on pooled semantic features. A sub-discriminator evaluates the progressive multi-stage outputs using classification and projection heads, optimized via a semantic alignment loss. We train the framework using a two-stage strategy: Stage 1 pre-trains the generators, and Stage 2 trains a semantically aligned classifier on source and synthesized target-domain patches using Sharpness-Aware Minimization (SAM). The proposed framework consistently outperforms existing domain generalization methods on multiple benchmark HSI datasets. By grounding progressive generation in vision-language semantics, our method improves cross-domain representation learning and enhances generalization for hyperspectral image classification.
\end{abstract}
